\vspace{-2cm} Inference on the Effects of Observed Features \\ in Latent Space Models for Networks

### 2018-03-12 manuscript_rearranged.tex
updates

And in sociology \citet[p. 157]{spillane2017elephant},

 \begin{quote} ``the models include latent space positions that implicitly control for structural aspects of networks, such as reciprocity and centrality, allowing for estimation of covariate effects (Hoff, Raftery, and Handcock 2002).''
 \end{quote}

 and  \citet[p. 107]{berlusconi2017determinants}

 \begin{quote} ``Higher order dependencies can be taken into account by assigning each country a position in a ?social space,? based on its (unobserved) characteristics and the presence of other ties. Conditional independence between observations can then subsequently be assumed, given the countries? latent position within the social space (Hoff et al., 2002).''
 \end{quote}

@@

For simplicity and computational efficiency we consider a
unidimensional latent space. We expect our results to generalize to
higher dimensions and other variants of the LSM since each variant
involves the use of unconstrained latent variables to represent
unmeasured network structure. To generate an observed covariate that
exhibits a controllable degree of colinearity with the latent network
structure we follow a three-step process. First, we simulate
unidimensional positions for each node, drawing from a normal
distribution, and calculate the Euclidean distance $\mathbf{d}$
between each pair of positions. Second, given a target covariance
matrix ($\Sigma$) among the covariates and distances, $\langle
\mathbf{x}, \mathbf{d} \rangle$, we derive the conditional mean vector
and covariance, assuming that $\mathbf{x}$ has a normal distribution
given $\mathbf{d}$ (see \cite[pp. 116--117]{eaton1983} for the
conditional normal derivation). Third, we simulate $x$ as a normal
random variable with the respective conditional means and
covariance. Finally we standardize $x$ to have zero mean and unit
variance. To generate the covariance matrix $\Sigma$, which controls
the dependence between the omitted network structure and the observed
covariate $\mathbf{x}$ we utilize the C-vine method of
\cite{lewandowski2009generating}.

We consider three exponential family distributions from which the
adjacency matrix entries are drawn: Gaussian, Bernoilli, and
Poisson. We also vary the size of the network under study, considering
networks with $n = 25, 50,$ and $100$ nodes. To vary the degree of
confounding attributable to the latent positions, we consider three
values of the collinearity parameter $\eta$, $1$, $100$, and
$1,000,000$, where $1$ corresponds to a standard uniform distribution
of the correlation (i.e., moderate collinearity) between the observed
covariate and the latent distances, and $1,000,000$ to independence
between the observed covariate and the latent distances. See Figure
\ref{fig:vine} for the distribution of the absolute value of the
correlation generated at each value of $\eta$.

@@

For each combination of simulation conditions we evaluate the mean
square prediction error of the model on new edges drawn condtional on
$\mathbf{x}$ and $\mathbf{d}$, the bias of the estimated coefficient
for the measured covariate, and the Type-1 and 2 error
rates. Inference regarding Credible intervals for $\beta$ are defined
by using the region of highest posterior density which covers $95\%$
of the marginal posterior distribution
\cite{turkkan1993computation}. We consider $\beta$ to be statistically
significant at the 0.05 level if the 95\% credible interval excludes
0.

>>>>

@@

+For simplicity and computational efficiency we consider a unidimensional latent space, that is, we create a single unmeasured variable that can be represented using the Euclidean parameterization of the LSM. We expect our results to generalize to higher dimensions and other variants of the LSM since each variant involves the use of unconstrained latent variables to represent unmeasured network structure. To generate an observed covariate that exhibits a controllable degree of colinearity with the latent network structure we follow a four-step process. First, we simulate unidimensional positions for each node, drawing from a standard normal distribution, and calculate the Euclidean distance $\mathbf{d}$ between each pair of positions. Second, we generate a target covariance matrix ($\Sigma$) using the $c$-vine method of \cite{lewandowski2009generating} which generates a uniform draw from the space of positive-definite matrices. The $c$-vine method makes use of a parameter $\eta$, which when equal to $1$ generates draws from uniformly from the space of positive-definite matrices, making the correlation between the latent positions and the measured covariate 0 in expectation. As $\eta$ increases above $1$ draws of $\Sigma$ have increasingly weak correlation\footnote{We show the distribution of correlations under each value of $\eta$ used in our simulation in Figure \ref{fig:vine}.}. Third, given the target covariance matrix ($\Sigma$) among the covariates and distances, $\langle \mathbf{x}, \mathbf{d} \rangle$, we derive the conditional mean vector and covariance, assuming that $\mathbf{x}$ has a normal distribution given $\mathbf{d}$ (see \cite[pp. 116--117]{eaton1983} for the conditional normal derivation).
+Fourth, we simulate $x$ as a normal random variable with the respective conditional means and covariance. Finally we standardize $x$ to have zero mean and unit variance to ensure that the scale of estimation errors is comparable across simulations.

We consider three exponential family distributions from which the adjacency matrix entries are drawn: Gaussian, Bernoilli, and Poisson. We also vary the size of the network under study, considering networks with $n = 25, 50,$ and $100$ nodes. To vary the degree of confounding attributable to the latent positions, we consider three values of the collinearity parameter $\eta$: $1$, $100$, and $1,000,000$. Here  where $\eta = 1$ gives a matrix where the correlation is in expectation $.5$, $\eta = 100$ results in moderate collinearity between the observed covariate and the latent distances, and $\eta = 1,000,000$ corresponds to independence between the observed covariate and the latent distances. See Figure \ref{fig:vine} for the distribution of the absolute value of the correlation generated at each value of $\eta$.

@@

For each combination of simulation conditions we evaluate the mean
square prediction error of the model on new edges drawn condtional on
$\mathbf{x}$ and $\mathbf{d}$: $\frac{1}{n}\sum_{i=1}^{n}(y_{i} - \hat{y}_i)^2$, the bias of the estimated coefficient
for the measured covariate $\frac{1}{n} \sum_{i=1}^n (\beta - \hat{beta})$, and the Type-1 and 2 error
rates. These error rates are estimated by computing coverage: $\frac{1}{m} \sum_{i=1}^m\mathbb{I}(\beta \in \text{HPD}(\hat{\beta}_i))$, where $\text{HPD}(\hat{beta})$ is the region of highest posterior density which covers $95\%$ of the marginal posterior distribution and $m$ is the number of samples from said posterior \cite{turkkan1993computation}. When in the simulation $\beta = 0$, $1 - P(\text{coverage})$ gives the type-1 error rate, and $P(\text{coverage})$ givesthe type-2 error rate when $\beta = 1$. Hence we consider $\beta$ to be statistically
significant at the 0.05 level if the 95\% credible interval excludes 0.
